### 2015-04-08 pract3/paper/paper.tex
first paper

\documentclass[11pt,a4paper]{article}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{graphicx}
\author{Bastiaan Weijers \\ 3256669 \and Colin Smits \\ 4075390}
\title{Evaluation of a Body Photo Browser}
\begin{document}
\maketitle
\abstract{Here comes the abstract}\\
\textbf{keywords:}

\noindent\makebox[\linewidth]{\rule{\textwidth}{0.4pt}}
\section{Introduction}
\section{System Evaluation}
\section{User Evaluation}
\section{Discussion}

\thebibliography{99}

\end{document}